---
title: "World Happiness Report Data Clean"
author: "Anish Nathu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
```{r}
library(dplyr)
library(tidyverse)
```


# Datasets
```{r}
WHR2015 <- read.csv("Data/2015.csv",sep=',', header = TRUE)
head(WHR2015)
```

```{r}
WHR2016 <- read.csv("Data/2016.csv",sep=',', header = TRUE)
head(WHR2016)
```

```{r}
WHR2017 <- read.csv("Data/2017.csv",sep=',', header = TRUE)
head(WHR2017)
```

```{r}
WHR2018 <- read.csv("Data/2018.csv",sep=',', header = TRUE)
head(WHR2018)
```

```{r}
WHR2019 <- read.csv("Data/2019.csv",sep=',', header = TRUE)
head(WHR2019)
```
# Structure
```{r}
str(WHR2015)
```

```{r}
str(WHR2016)
```

```{r}
str(WHR2017)
```

```{r}
str(WHR2018)
```

```{r}
str(WHR2019)
```


# Summary
```{r}
summary(WHR2015)
```

```{r}
summary(WHR2016)
```

```{r}
summary(WHR2017)
```

```{r}
summary(WHR2018)
```

```{r}
summary(WHR2019)
```

# Dimensions
```{r}
dim(WHR2015)
```

```{r}
dim(WHR2016)
```

```{r}
dim(WHR2017)
```

```{r}
dim(WHR2018)
```

```{r}
dim(WHR2019)
```

# Column Names
```{r}
colnames(WHR2015)
```

```{r}
colnames(WHR2016)
```

```{r}
colnames(WHR2017)
```

```{r}
colnames(WHR2018)
```

```{r}
colnames(WHR2019)
```

# Adding, Removing and Renaming Columns to the Dataframe of all years
```{r}
clean_WHR2015 <- WHR2015 %>%
  rename('Economy' = "Economy..GDP.per.Capita.",
         'Rank' = "Happiness.Rank",
         'Score' = "Happiness.Score",
         'Economy' = "Economy..GDP.per.Capita.",
         'Health' = "Health..Life.Expectancy.",
         'Government' = 'Trust..Government.Corruption.',
         'Dystopia' = "Dystopia.Residual") %>%
  add_column('Year' = 2015) %>%
  select(-'Standard.Error', -'Rank', -'Dystopia')

head(clean_WHR2015)
```

```{r}
clean_WHR2016 <- WHR2016 %>%
  rename('Economy' = "Economy..GDP.per.Capita.",
         'Rank' = "Happiness.Rank",
         'Score' = "Happiness.Score",
         'Economy' = "Economy..GDP.per.Capita.",
         'Health' = "Health..Life.Expectancy.",
         'Government' = 'Trust..Government.Corruption.',
         'Dystopia' = "Dystopia.Residual") %>%
  add_column('Year' = 2016) %>%
  select(-'Lower.Confidence.Interval', -'Upper.Confidence.Interval', -'Rank', -'Dystopia')

head(clean_WHR2016)
```

```{r}
clean_WHR2017 <- WHR2017 %>%
  rename('Economy' = "Economy..GDP.per.Capita.",
         'Rank' = "Happiness.Rank",
         'Score' = "Happiness.Score",
         'Economy' = "Economy..GDP.per.Capita.",
         'Health' = "Health..Life.Expectancy.",
         'Government' = 'Trust..Government.Corruption.',
         'Dystopia' = "Dystopia.Residual") %>%
  add_column('Year' = 2017, 'Region' = NA) %>%
  select(-'Whisker.high', -'Whisker.low',-'Rank', -'Dystopia')

head(clean_WHR2017)
```

```{r}
clean_WHR2018 <- WHR2018 %>%
  rename ('Rank' = "Overall.rank",
          'Country' = "Country.or.region",       
          'Economy' = "GDP.per.capita",
          'Family' = 'Social.support',
          'Health' = "Healthy.life.expectancy",
          'Freedom' = "Freedom.to.make.life.choices",
          'Government' = "Perceptions.of.corruption") %>%
  add_column ('Year' = 2018, 'Region' = NA ) %>%
  select (-'Rank')

head(clean_WHR2018)
```

```{r}
clean_WHR2019 <- WHR2019 %>%
  rename('Rank' = "Overall.rank",
         'Country' = "Country.or.region",       
         'Economy' = "GDP.per.capita",
         'Family' = 'Social.support',
         'Health' = "Healthy.life.expectancy",
         'Freedom' = "Freedom.to.make.life.choices",
         'Government' = "Perceptions.of.corruption") %>%
  add_column ('Year' = 2019, 'Region' = NA) %>%
  select (-'Rank')

head(clean_WHR2019)
```

# Merging the dataframes together
```{r}
WHR_Cleaned <- rbind(clean_WHR2015, clean_WHR2016, clean_WHR2017, clean_WHR2018, clean_WHR2019) 
head(WHR_Cleaned)
```

# Structure of Cleaned Dataframe
```{r}
str(WHR_Cleaned)
```
# Changing the Trust in Govt to numeric data type
```{r}
WHR_Cleaned$Government<-as.numeric(WHR_Cleaned$Government)
sapply(WHR_Cleaned, mode)
```

# Summary of Cleaned Dataframe
```{r}
summary(WHR_Cleaned)
```

# Dimensions of Cleaned Dataframe
```{r}
dim(WHR_Cleaned)
```

# Column Names of Cleaned Dataframe
```{r}
colnames(WHR_Cleaned)
```

# Changing the Order of Cleaned Dataframe
```{r}
order <- c("Country","Region","Year", "Score","Economy",
           "Family","Health", "Freedom","Government","Generosity")
WHR_Cleaned <- WHR_Cleaned[, order]
head(WHR_Cleaned)
```

# Difference in countries names over the years
```{r}
unique(sort(WHR_Cleaned$Country))
```

# All the different Regions in the dataframe
```{r}
unique(sort(WHR_Cleaned$Region))
```

# Changing countries names
```{r}
WHR_Cleaned$Country[WHR_Cleaned$Country=="Hong Kong S.A.R., China"] <- "Hong Kong"

WHR_Cleaned$Country[WHR_Cleaned$Country=="Macedonia"] <- "North Macedonia"

WHR_Cleaned$Country[WHR_Cleaned$Country=="North Cyprus"] <- "Northern Cyprus"

WHR_Cleaned$Country[WHR_Cleaned$Country=="Somaliland region"] <- "Somalia"
WHR_Cleaned$Country[WHR_Cleaned$Country=="Somaliland Region"] <- "Somalia"

WHR_Cleaned$Country[WHR_Cleaned$Country=="Taiwan Province of China"] <- "Taiwan"

WHR_Cleaned$Country[WHR_Cleaned$Country=="Trinidad & Tobago"] <- "Trinidad and Tobago"
```


# Number of NA's in the Cleaned Dataframe
```{r}
colSums(is.na(WHR_Cleaned))
```

# Fixing all the NA's values
# First is the Government missing value
```{r}
WHR_Cleaned[is.na(WHR_Cleaned$Government),]
```

```{r}
filter(WHR_Cleaned, Country == "United Arab Emirates")
```
# Imputed by using the mean of Government column.
```{r}
WHR_Cleaned$Government[is.na(WHR_Cleaned$Government)] <- mean(WHR_Cleaned$Government, na.rm = T)  
filter(WHR_Cleaned, Country == "United Arab Emirates")
```

# Lastly adding all the missing values for the Region Column
# First sort by Country and Year making it alphabetical and ordered year
```{r}
WHR_Cleaned <- arrange(WHR_Cleaned, Country, Year)
WHR_Cleaned
```

```{r}
WHR_Cleaned <- WHR_Cleaned %>%
  group_by(Country) %>%
  fill(Region, .direction = c('updown'))
WHR_Cleaned
```

# Checking the Number of NA's in the Dataframe after Cleaning
```{r}
colSums(is.na(WHR_Cleaned))
```

# Finding and Fixing the last NA value
```{r}
WHR_Cleaned[is.na(WHR_Cleaned$Region),]
```

# Imputed by this: https://data.worldbank.org/country/ZG
```{r}
WHR_Cleaned$Region[is.na(WHR_Cleaned$Region)] <- "Sub-Saharan Africa"
filter(WHR_Cleaned, Country == "Gambia")
```

# Final Check of the Number of NA's in the Dataframe after Cleaning
```{r}
colSums(is.na(WHR_Cleaned))
```

# Writing the cleaned csv for World Happiness Report for the years of 2015 to 2019
```{r}
write.csv(WHR_Cleaned, file = 'Data/WHR_Cleaned_Data.csv', row.names = FALSE)
```